---
title: "Fisher Lab SlakeItEasy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example image processing workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette outlines a typical workflow using SlakeItEasy to quantify soil aggregate stability from images. See the package [GitHub repository](https://github.com/Soil-Health-Institute/SlakeItEasy/) for instructions on package installation and required file structure and naming conventions. If you want to analyze soils in the field and only need to process one soil sample at a time, check out the Slakes app for Android or iPhone.

A brief note for users without prior R experience: SlakeItEasy is a set of tools (called *functions*) designed to make it easier to quantify the expansion of soils submerged in water. We make a few assumptions about how images were collected and the settings for image processing in the workflow illustrated here and in the default function arguments. To learn more about what each function does and the settings that can be changed, type `?functionname` in the console.

##download EBimage first time
if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("EBImage")

##install slake it easy first time
# install {remotes} if missing
if (!require(remotes)) {
install.packages('remotes')
}
# install latest version of {SlakeItEasy} from GitHub repository
remotes::install_github('Soil-Health-Institute/SlakeItEasy')


## START HERE if you have everything installed
## Load required packages

```{r setup, message=FALSE}
library(EBImage)
library(dplyr)
library(SlakeItEasy)
```

## this just sets formatting, so not necessary to run

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts = list(width.cutoff = 60), tidy = TRUE
)
```



## Set paths to images and output directory

SlakeItEasy is designed to analyze batches of images. We need to tell R where on the computer the image batch is saved and where we want to save files related to image processing and our final aggregate stability data.

The SlakeItEasy package provides the function `set_paths()` for users that prefer to use the file browser to specify locations of images and file output. YOU WILL HAVE ERRORS IF YOU DON'T SET AN OUTPUT DIRECTORY. For output, select a general directory where results for multiple batches can be saved.

```{r, eval=FALSE}
paths <- set_paths()
```

Alternatively, you can specify `paths` non-interactively by defining a named list with the image directory, output directory, and batch name.

```{r, eval=F}
paths <- list(image_dir = '~/Documents/r/Slake/CSC/Summer2022/S22W1005/S22W1005-RP2',
              output_dir = dir.create('~/Documents/r/Slake/CSC/Summer2022/S22W1005/S22W1005-RP2/output'),
              batch_name = 'bafbatch')
```

```{r, include=FALSE}
paths <- list(image_dir = system.file("images/", package="SlakeItEasy"),
              output_dir = tempdir(),
              batch_name = paste0('batch_', format(Sys.Date(), '%Y%m%d')))
```

## Set up directories for output

Create a batch-specific directory for processing outputs. By default, `dir_setup` will also create subdirectories that we will use later to sort images.

```{r}

dirs_for_flagged_imgs <- dir_setup(paths$output_dir,  return_paths = T)

# extract paths to subdirectories for images that need to be processed interactively and for images that cannot be analyzed (requiring samples be rerun)

reprocess_dir <- dirs_for_flagged_imgs['to_reprocess']
unusable_dir <- dirs_for_flagged_imgs['unusable']
```

## Inspect image metadata

SlakeItEasy extracts image attributes from file names, though the document says it uses EXIF metadata, it actually uses \image file names. Fisher Lab had a different camera format than slakes developers, so we created this block of commands to rename files to fit the Slake It Easy file name format. These steps rename files from Canon EOS T3 by pulling the date and time from the image EXIF and renaming the file with the date and time. 

##Canon EOS T3 metadata format for Fisher Lab: <xmp:CreateDate>2025-02-21T14:13:44</xmp:CreateDate>
## also formatted as 2025-02-21 14:13:44.013 

# To rename images from EXIF data, Load required packages
install.packages("exifr")  # If not already installed
library(exifr)

# image_dir was set above. 
# Get the list of all .jpg files in the root directory and subdirectories and view them. 
#this set of steps is not strictly necessary
image_files <- list.files(image_dir, pattern = "\\.jpg$", full.names = TRUE, recursive = TRUE, ignore.case = TRUE)
image_files

# Rename images in the defined directory from EXIF data 
if (length(image_files) == 0) {
  cat("No images found in the specified directory.\n")
} else {
  # Loop through each image file and rename based on CreateDate
  for (image_path in image_files) {

    # Extract EXIF metadata
    metadata <- read_exif(image_path, tags = "CreateDate")

    # Check if CreateDate exists
    if (!is.na(metadata$CreateDate)) {

      # Format CreateDate into IMG_YYYYMMDD_HHMMSS.jpg
      new_name <- format(as.POSIXct(metadata$CreateDate, format="%Y:%m:%d %H:%M:%S"), "IMG_%Y%m%d_%H%M%S.jpg")

      # Keep the file in the same directory
      image_dir <- dirname(image_path)  # Get the current folder of the image
      new_path <- file.path(image_dir, new_name)  # Define new full path

      # Rename the file if the new name doesn't already exist
      if (!file.exists(new_path)) {
        file.rename(image_path, new_path)
        cat("Renamed:", image_path, "->", new_name, "\n")
      } else {
        cat("Skipped:", image_path, "(Target filename already exists)\n")
      }
    } else {
      cat("No CreateDate found for:", image_path, "\n")
    }
  }
}

```{r}
metadata <- get_metadata(paths$image_dir, filename_prefix = 'IMG_')
```

Images in the directory specified by `paths$image_dir` are assumed to be arranged by replicate: that is, an individual set of aggregates that were photographed in the air-dry state, immediately after submersion, and after a specified time in water  (by default, ten minutes ±30 seconds). Use `check_replicates` to identify image sets that deviate from these expectations.

```{r}
metadata_qaqc <- check_replicates(metadata,  final_img_time_min = 10, final_img_tol_sec = 30, n_images_max = 3)

# save metadata summary to output directory
write.csv(metadata_qaqc$m, file.path(paths$output_dir, 'qaqc_log.csv'), row.names = F)

metadata_qaqc$usable
metadata_qaqc$missing_imgs
metadata_qaqc$extra_imgs
metadata_qaqc$wrong_final_time

# metadata[metadata$Directory %in% metadata_qaqc$wrong_final_time,]

# inspect replicates with extra images
if (length(metadata_qaqc$extra_imgs) > 0) {
  opendirs(metadata_qaqc$extra_imgs)
}
```

## Process individual images

The suitability of default function arguments and need for alternative values can be evaluated by running `mask_image_circular()` and/or `area_from_image()`. If the sample container (Petri dish) is off-center in the image, modify the mask offsets (`h_offset` and `v_offset`). If the mask needs to be resized, modify the relative diameter of the mask (`d`).

```{r, eval = F}
img <- read_to_portrait(with(metadata[1,], paste0(Directory, '/', FileName)))
img_masked <- mask_image_circular(img, interactive = T, h_offset = 0)
plot(img_masked)

test <- area_from_image(with(metadata[1,], paste0(Directory, '/', FileName)), circular_mask = T, d = 0.9, interactive = T)
```

## Automated batch processing with a circular crop

```{r, eval = F}
results <- batch_process(dir_vec = to_analyze, outdir = paths$output_dir, filename_prefix = 'IMG_', match_resolution = T, circular_mask = T, parallel = T)
```

## Inspect classifications and sort suboptimal images

```{r, eval = F}
reprocess_dir <- dirs_for_flagged_imgs['to_reprocess']
unusable_dir <- dirs_for_flagged_imgs['unusable']
opendirs(c(unusable_dir, reprocess_dir, paste0(paths$output_dir, '/images_false_color')))
```

## Prepare for manual processing

```{r, eval = F}
images_to_reprocess <- list.files(reprocess_dir)

dir_vec2 <- unique(sapply(strsplit(images_to_reprocess, '_x_'), function(x) paste(x[1:(length(x) - 1)], collapse = '/')))
```
